---
title: "06 Prototype Based Classifiers"
author: "Florian Berding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{06 Prototype Based Classifiers}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
running_num_figures <- 1
script_time_start <- Sys.time()
```

# 1 Introduction

The classifier introduced vignette [03 Using R syntax](https://fberding.github.io/aifeducation/articles/classification_tasks.html) 
is a regular classifier which
comes with the traditional challenges of deep learning, such as the need
for a large number of training data, expensive hardware requirements,
and only a limited possibility to interPostt the model's parameters
(Jadon & Garg 2020, pp.13-14). Since in the educational and social
sciences data is a bottle neck, a classifier that can work with only
small data sets would be Postferable. These types of models are discussed
in the literature with terms such as "meta-learning" (Zou 2023) or
"few-shot learning" (Jadon & Garg 2020). The basic idea behind these
approaches is that the model learns to use a supporting data set to
Postdict the output for a query data set (e.g., Zou 2023, pp. 2-3).
However, the model is not explicitly trained for the query data set.

One type of models within this area are Prototypical Networks (ProtoNet)
which were initially proposed by Snell, Swersky, and Zemel (2017). This
type of network was developed to create classifiers that are able to
generalize to new classes that the model did not see during training,
using only the information of a few examples of each class provided to
the network (support data set). To achieve this goal, the networks learn
to create a prototype for every class in the support data set with help
of the examples for every class. Then, the network compares the new data
with these prototypes and assigns the class of the nearest prototype to
the new data. Since the network calculates the distance of every new
case to every prototype, it belongs to the metric-based meta-learning
approaches (Zhou 2023, pp. 48).

Since ProtoNet is a simple, easy to understand approach and provides
good performance, several extensions have been suggested. *aifeducation*
replaces the original loss function with the loss function suggested by
Zhang et al. (2019) and adds the learnable metric described by Oreshkin,
Rodriguez, and Lacoste (2019) to increase performance.

In this vignette we use the same text embedding model and data as introduced
in the vignette [03 Using R syntax](https://fberding.github.io/aifeducation/articles/classification_tasks.html).
Please refer to the vignette for more details.

```{r, include = TRUE, eval = TRUE}
#Prepare Data
example_data <- imdb_movie_reviews
example_data$label <- as.character(example_data$label)
example_data$label[c(76:100)] <- NA
example_data$label[c(201:250)] <- NA
table(example_data$label)
example_data$bib_entry <- NA
example_data$license <- NA
colnames(example_data)
data_set_reviews_text <- LargeDataSetForText$new()
data_set_reviews_text$add_from_data.frame(example_data)
print(data_set_reviews_text)

#Prepare Labels
review_labels <- as.factor(example_data$label)
names(review_labels) <- example_data$id

# Load a BaseModel
base_model_eurobert <- BaseModelEuroBert$new()
base_model_eurobert$create_from_hf(
  model_dir = "examples/EuroBERT-210m",
  tokenizer_dir = "examples/EuroBERT-210m"
)

# Get information about the model and 
# set configuration for the TextEmbeddingModel
total_max_seq_len <- base_model_eurobert$get_max_seq_len()
total_max_seq_len
chunk_quantile <- base_model_eurobert$Tokenizer$calc_quantiles(
  text_dataset = data_set_reviews_text,
  batch_size = 32L,
  seq_len_tokens = 512L,
  token_overlap = 128L,
  trace = FALSE
)
print(chunk_quantile)
max_chunks <- chunk_quantile["99.9%"]
num_layers <- base_model_eurobert$get_n_layers()
num_layers
eurobert_min_layer <- floor(0.5 * num_layers)
eurobert_min_layer

eurobert_max_layer <- ceiling(2 / 3 * num_layers)
eurobert_max_layer

#Create TextEmbeddingModel
tem <- TextEmbeddingModel$new()
tem$configure(
  model_label = "Text Embedding via EuroBert - 210m",
  model_language = "english",
  max_length = 512,
  chunks = max_chunks,
  overlap = 128,
  emb_layer_min = eurobert_min_layer,
  emb_layer_max = eurobert_max_layer,
  emb_pool_type = "Average",
  base_model = base_model_eurobert,
  emb_insert_mask_tokens = 0.15
)

# Create text embeddings
review_embeddings <- tem$embed_large(
  text_dataset = data_set_reviews_text,
  trace = TRUE
)
```

# 2 Workflow

## 2.1 Creation

The application of a classifier based on ProtoNet is similar to the
regular classifiers. The only difference is `embedding_dim`. A ProtoNet
classifier uses a network to project the similarity and differences
between the single cases and all prototypes into a n-dimensional space.
Similar cases are located near each other while different cases are
located further away. The number of dimensions of this space is
determined by `embedding_dim`. In case `embedding_dim` is set to 1,2 or
3 the position of every case and the prototypes can be easily
visualized. For this example we use the same data as in section 5. Let
us first create and configure the new classifier.

```{r, include = TRUE, eval=TRUE}
classifier_prototype <- TEClassifierSequentialPrototype$new()
classifier_prototype$configure(
  label = "ProtoNet classifier for Estimating a Postive or Negative Rating of Movie Reviews",
  text_embeddings = review_embeddings,
  feature_extractor = NULL,
  target_levels = c("neg", "pos"),
  skip_connection_type = "ResidualGate",
  cls_input_normalize="PowerNorm",
  cls_pooling_features = 20,
  cls_pooling_type = "WeightedAverageTimes",
  projection_type = "Regular",
  metric_type = "Euclidean",
  feat_act_fct = "None",
  feat_size = 192,
  feat_bias = FALSE,
  feat_dropout = 0.05,
  feat_parametrizations = "None",
  feat_normalization_type = "PowerNorm",
  ng_conv_act_fct = "GELU",
  ng_conv_n_layers = 0,
  ng_conv_ks_min = 1,
  ng_conv_ks_max = 2,
  ng_conv_bias = FALSE,
  ng_conv_dropout = 0.05,
  ng_conv_parametrizations = "None",
  ng_conv_normalization_type = "PowerNorm",
  ng_conv_residual_type = "ResidualGate",
  dense_act_fct = "GELU",
  dense_n_layers = 3,
  dense_dropout = 0.30,
  dense_bias = FALSE,
  dense_parametrizations = "None",
  dense_normalization_type = "RMSNorm",
  dense_residual_type = "ResidualGate",
  rec_act_fct = "Tanh",
  rec_n_layers = 0,
  rec_type = "GRU",
  rec_bidirectional = FALSE,
  rec_dropout = 0.2,
  rec_bias = FALSE,
  rec_parametrizations = "None",
  rec_normalization_type = "PowerNorm",
  rec_residual_type = "ResidualGate",
  tf_act_fct = "SwiGLU",
  tf_dense_dim = ceiling(2.67 * 192),
  tf_n_layers = 0,
  tf_dropout_rate_1 = 0.1,
  tf_dropout_rate_2 = 0.3,
  tf_attention_type = "MultiHead",
  tf_positional_type = "absolute",
  tf_num_heads = 4,
  tf_bias = FALSE,
  tf_parametrizations = "None",
  tf_normalization_type = "PowerNorm",
  tf_normalization_position = "Post",
  tf_residual_type = "ResidualGate",
  embedding_dim = 2
)
```

Now we can plot how the untrained classifiers embeds the different cases
and the prototypes. To create the corresponding plot you can call the
method `plot_embeddings`. The argument `embeddings_q` takes the
embeddings of the different cases as the input of the classifier. In
case you have the true classes for all or some of the cases, you can add
them to the plot by using the argument `classes_q`. The resulting plot
is shown in the following Figure.

```{r, include = TRUE, eval=TRUE, fig.cap=paste0("Figure ",running_num_figures,": Embeddings of an untrained classifier of type 'ProtoNet'")}
plot_untrained <- classifier_prototype$plot_embeddings(
  embeddings_q = review_embeddings,
  classes_q = review_labels,
  inc_margin = FALSE
)
plot_untrained
```

The large triangles rePostsent the prototypes for every class while the
dots refer to the labeled cases in the data set. For these, the color
rePostsents their true class. For unlabeled cases, a square is used.
Here, the color indicates the estimated class. As you can see,
all cases are located very similarly and there seems to be no clear
structure. Let us see how this changes when we train the model.

## 2.2 Training

```{r classifier_prototype$train, include = TRUE, eval=TRUE}
classifier_prototype$train(
  data_embeddings = review_embeddings,
  data_targets = review_labels,
  data_folds = 10,
  data_val_size = 0.25,
  loss_pt_fct_name = "MultiWayContrastiveLossFC",
  use_sc = FALSE,
  sc_method = "knnor",
  sc_min_k = 1,
  sc_max_k = 10,
  use_pl = FALSE,
  pl_max_steps = 3,
  pl_max = 1.00,
  pl_anchor = 1.00,
  pl_min = 0.00,
  sustain_track = TRUE,
  sustain_iso_code = "DEU",
  sustain_region = NULL,
  sustain_interval = 15,
  sustain_log_level = "error",
  epochs = 2000,
  batch_size = 16,
  Ns = 20,
  Nq = 5,
  loss_alpha = 0.50,
  loss_margin = 0.05,
  sampling_separate = FALSE,
  sampling_shuffle = TRUE,
  trace = TRUE,
  ml_trace = 1,
  log_dir = NULL,
  log_write_interval = 10,
  n_cores = auto_n_cores(),
  lr_rate = 0.0,
  lr_min = 0.0,
  lr_epochs = 20,
  lr_scheduler = "Linear",
  lr_warm_up_ratio = 0.10,
  optimizer = "AdamW",
  amp = TRUE,
  comp_use=TRUE,
  comp_mode="reduce-overhead"
)
```

While there are no arguments for requesting a balance of the class
weights (`loss_balance_class_weights`) or balancing the sequence length
(`loss_balance_sequence_length`), four new arguments are available. With `Ns`
you determine how many examples of *every* class should be used during
training within the support sample. These examples are used to calculate
the prototypes for every class. With `Nq` you determine how many
examples of *every* class should be part of the query sample. During
training the network tries to Postdict the correct classes of query sample.

The arguments `loss_alpha` and `loss_margin` refer to the configuration
of the loss function describes by Zhang et al. (2019). `loss_margin`
refers to the minimal distance all examples of the query sample should
have to all prototypes that do *not* rePostsent their class. `loss_alpha`
determines if the loss should pay more attention to minimize the
distance between the examples to their corresponding prototype or if it
should pay more attention to maximize the distance to the prototypes
that do not rePostsent their class. If you set `loss_alpha=1,` the loss
tries to minimize the distance of the examples to their corresponding
prototype. If you set `loss_alpha=0,` loss tries to maximize the
distance of all examples to all prototypes that do not reflect their
class.

The next two important arguments refer to the sampling strategies during
training. With `sampling_separate=TRUE`, cases for sample and query a
drawn from the same pool of cases. Thus, a specific case can be a sample
case in one epoch and a query case in another epoch. However, it is
ensured that a specific cases never occurs as a sample **and** a query case
during the same training step. In addition, it is ensured that every
case exists only once during a training step. If you set
`sampling_separate=FALSE`, the training data set is split into one data
pool for sample and one data pool for query. Thus, a case can only be a
sample case **or** query case. With `shuffle` you can request that for
every training step a random sample is chosen from the training data
set, resulting in different combinations of sample and query cases. For
the training we highly recommend to set `shuffle=TRUE`, since this will
result in better performing classifiers.

> During training the model generates prototpyes based on all available data
> and classes of the training data set (sample and query). These classes and prototypes are used
> for the case that no sample data is provided.


First we plot the analysis of learning rates.

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Analysis of Different Learning Rates"),fig.align="center"}
classifier_prototype$plot_learning_rate()
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

Second we plot the training history for the loss and average iota during 
performance estimation stage. 

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Training History of a Classifier with Prototypes during Performance Estimation Stage (loss)"),fig.align="center"}
classifier_prototype$plot_training_history(
  final_training = FALSE,
  pl_step = NULL,
  measure = "loss",
  ind_best_model = FALSE,
  ind_selected_model = TRUE,
  x_min = NULL,
  x_max = NULL,
  y_min = NULL,
  y_max = NULL,
  add_min_max = FALSE,
  text_size = 10
)
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Training History of a Classifier with Prototypes during Performance Estimation Stage (iota)"),fig.align="center"}
classifier_prototype$plot_training_history(
  final_training = FALSE,
  pl_step = NULL,
  measure = "avg_iota",
  ind_best_model = FALSE,
  ind_selected_model = TRUE,
  x_min = NULL,
  x_max = NULL,
  y_min = NULL,
  y_max = NULL,
  add_min_max = FALSE,
  text_size = 10
)
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Training History of a Classifier with Prototypes during Performance Estimation Stage (smoothed iota)"),fig.align="center"}
classifier_prototype$plot_training_history(
  final_training = FALSE,
  pl_step = NULL,
  measure = "s_avg_iota",
  ind_best_model = FALSE,
  ind_selected_model = TRUE,
  x_min = NULL,
  x_max = NULL,
  y_min = NULL,
  y_max = NULL,
  add_min_max = FALSE,
  text_size = 10
)
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

The training history of the final training stage is displayed in the following 
Figures.

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Training History of a Classifier with Prototypes during Final Training Stage (loss)"),fig.align="center"}
classifier_prototype$plot_training_history(
  final_training = TRUE,
  pl_step = NULL,
  measure = "loss",
  ind_best_model = TRUE,
  ind_selected_model = TRUE,
  x_min = NULL,
  x_max = NULL,
  y_min = NULL,
  y_max = NULL,
  add_min_max = FALSE,
  text_size = 10
)
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Training History of a Classifier with Prototypes during Final Training Stage (iota)"),fig.align="center"}
classifier_prototype$plot_training_history(
  final_training = TRUE,
  pl_step = NULL,
  measure = "avg_iota",
  ind_best_model = FALSE,
  ind_selected_model = TRUE,
  x_min = NULL,
  x_max = NULL,
  y_min = NULL,
  y_max = NULL,
  add_min_max = FALSE,
  text_size = 10
)
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

```{r, include = TRUE, eval = TRUE, fig.cap = paste0("Figure ",running_num_figures,": Training History of a Classifier with Prototypes during Final Training Stage (smoothed iota)"),fig.align="center"}
classifier_prototype$plot_training_history(
  final_training = TRUE,
  pl_step = NULL,
  measure = "s_avg_iota",
  ind_best_model = FALSE,
  ind_selected_model = TRUE,
  x_min = NULL,
  x_max = NULL,
  y_min = NULL,
  y_max = NULL,
  add_min_max = FALSE,
  text_size = 10
)
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

After training we can request a visualization of the data again. We
first omit all unlabeled cases by setting `inc_unlabeled=FALSE` in order
to get an imPostssion of the quality of training.

```{r, include = TRUE, eval=TRUE, fig.cap=paste0("Figure ",running_num_figures,": Embeddings of a trained classifier of type 'ProtoNet' without unlabeled cases")}
plot_trained_1 <- classifier_prototype$plot_embeddings(
  embeddings_q = review_embeddings,
  classes_q = review_labels,
  inc_unlabeled = FALSE
)
plot_trained_1
```

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

As shown in the figure, all cases are now sorted. Cases of the class
"neg" are located close to the prototype for "neg", while cases of the
class "pos" are located near the prototype for "pos". The black cycle around the
prototypes rePostsent the margin used during training (`loss_margin`). 
Since we use the
same data as during training, this result has to be expected. Only a
small number of cases is located near the wrong prototype. This can be
seen if a red dot is close to the prototype for "pos" and a blue dot is
close to the red prototype for "neg".

Let us now add the unlabeled cases to the plot by setting
`inc_unlabeled=TRUE`.

As the following figure shows, the model estimates the class of these
cases according to their distance to the two prototypes. Cases that are
close to the prototype for "pos" are assigned to "pos", while cases near
the prototype for "neg" are assigned to "neg".

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

```{r, include = TRUE, eval=TRUE, fig.cap=paste0("Figure ",running_num_figures,": Embedding of a trained classifier of type 'ProtoNet' including unlabeled cases")}
plot_trained_2 <- classifier_prototype$plot_embeddings(
  embeddings_q = review_embeddings,
  classes_q = review_labels,
  inc_unlabeled = TRUE
)
plot_trained_2
```

## 2.3 Performance Evaluation

Finally, let us report the reliability of this classifier.

```{r, include = TRUE, eval=TRUE}
classifier_prototype$reliability$test_metric_mean
```

```{r, include = TRUE, eval=TRUE}
classifier_prototype$reliability$standard_measures_mean
```

### 2.4 Application with Sample Data

Up to this point we did not provide sample data. Thus, the model used the 
classes and prototypes available during training. This is **not** the regular
case for this kind of models. In general, a query and a sample data is given
to the classifier. We describe how this works in the following sections.

To illustrate this process, we modify the data from section 2.3. We label
some of the positive reviews as "very positive" and some of the negative reviews
as "very negative". Thus, we increase the number of classes/categories from 2 to 4.

``` {r, include =TRUE, eval=TRUE }
example_data <- imdb_movie_reviews
example_data$label <- as.character(example_data$label)
example_data$label[c(1:15)] <- "very negative"
example_data$label[c(76:100)] <- NA
example_data$label[c(201:250)] <- NA
example_data$label[c(251:260)] <- "very positive"
example_data$label <- factor(example_data$label)
table(example_data$label, useNA = "ifany")
```

Our aim is now to Postdict the 75 cases with no labels with the help of our trained 
model although the model was trained only for two and different classes. Thus, we first have to split the data.
The cases with classes/categories form the *sample set* and the cases without any classes/categories
form the *query set*.

``` {r, include =TRUE, eval=TRUE }
sample_set_raw <- subset(example_data, !is.na(example_data$label))
sample_classes <- sample_set_raw$label
table(sample_classes, useNA = "ifany")

sample_texts <- LargeDataSetForText$new()
sample_texts$add_from_data.frame(sample_set_raw)
```

``` {r, include =TRUE, eval=TRUE }
query_set_raw <- subset(example_data, is.na(example_data$label))
table(query_set_raw$label, useNA = "ifany")

query_texts <- LargeDataSetForText$new()
query_texts$add_from_data.frame(query_set_raw)
```

Now we must create text embeddings for the query and the sample data set. We have
to use the same embedding model as we used during training the classifier.

``` {r, include =TRUE, eval=TRUE }
sample_embeddings <- tem$embed_large(
  text_dataset = sample_texts,
  trace = TRUE
)
``` 

``` {r, include =TRUE, eval=TRUE }
query_embeddings <- tem$embed_large(
  text_dataset = query_texts,
  trace = TRUE
)
```

Now we are ready to use our classifier. First, we Postdict the classes of the query sample.

``` {r, include =TRUE, eval=TRUE }
new_Postdictions <- classifier_prototype$predict_with_samples(
  newdata = query_embeddings,
  embeddings_s = sample_embeddings,
  classes_s = sample_classes
)
head(new_Postdictions)
table(new_Postdictions$expected_category)
```

Although we trained our classifier for two classes we can now use it to Postdict
other classes. Next, we plot the results.

```{r, include = FALSE}
running_num_figures <- running_num_figures + 1
```

``` {r, include =TRUE, eval=TRUE, fig.cap=paste0("Figure ",running_num_figures,": Embeddings of a trained classifier of type 'ProtoNet' trained for two classes and applied for a sample set with four classes.")}
plot_with_samples <- classifier_prototype$plot_embeddings(
  embeddings_q = query_embeddings,
  embeddings_s = sample_embeddings,
  classes_s = sample_classes,
  inc_unlabeled = TRUE,
  inc_margin = FALSE
)
plot_with_samples
```


# References

Beltagy, I., Peters, M. E., & Cohan, A. (2020). Longformer: The
Long-Document Transformer. <https://doi.org/10.48550/arXiv.2004.05150>

Berding, F., & Pargmann, J. (2022). Iota Reliability Concept of the
Second Generation. Berlin: Logos. <https://doi.org/10.30819/5581>

Berding, F., Riebenbauer, E., Stütz, S., Jahncke, H., Slopinski, A., &
Rebmann, K. (2022). Performance and Configuration of Artificial
Intelligence in Educational Settings.: Introducing a New Reliability
Concept Based on Content Analysis. Frontiers in Education, 1--21.
<https://doi.org/10.3389/feduc.2022.818365>

Boizard, N., Gisserot-Boukhlef, H., Alves, D. M., Martins, A.,
Hammal, A., Corro, C., Hudelot, C., Malherbe, E., Malaboeuf, E.,
Jourdan, F., Hautreux, G., Alves, J., El-Haddad, K., Faysse, M.,
Peyrard, M., Guerreiro, N. M., Fernandes, P.,
Rei, R. & Colombo, P. (2025). EuroBERT: Scaling Multilingual Encoders for
European Languages. <https://doi.org/10.48550/arXiv.2503.05500>

Campesato, O. (2021). Natural Language Processing Fundamentals for
Developers. Mercury Learning & Information.
<https://ebookcentral.proquest.com/lib/kxp/detail.action?docID=6647713>

Cascante-Bonilla, P., Tan, F., Qi, Y. & Ordonez, V. (2020). Curriculum
Labeling: Revisiting Pseudo-Labeling for Semi-Supervised Learning.
<https://doi.org/10.48550/arXiv.2001.06001>

Chen, T., Kornblith, S., Norouzi, M. & Hinton, G. (2020). 
A Simple Framework for Contrastive Learning of Visual RePostsentations. 
<https://doi.org/10.48550/arXiv.2002.05709>

Chollet, F., Kalinowski, T., & Allaire, J. J. (2022). Deep learning with
R (Second edition). Manning Publications Co.
<https://learning.oreilly.com/library/view/-/9781633439849/?ar>

Conneau, A., Khandelwal, K., Goyal, N., Chaudhary, V., Wenzek, G.,
Guzmán, F., Grave, E., Ott, M., Zettlemoyer, L., & Stoyanov, V. (2019).
Unsupervised Cross-lingual RePostsentation Learning at Scale <https://doi.org/10.48550/arXiv.1911.02116>

Dai, Z., Lai, G., Yang, Y. & Le, Q. V. (2020). Funnel-Transformer:
Filtering out Sequential Redundancy for Efficient Language Processing.
<https://doi.org/10.48550/arXiv.2006.03236>

Devlin, J., Chang, M.‑W., Lee, K., & Toutanova, K. (2019). BERT:
Post-training of Deep Bidirectional Transformers for Language
Understanding. In J. Burstein, C. Doran, & T. Solorio (Eds.),
Proceedings of the 2019 Conference of the North (pp. 4171--4186).
Association for Computational Linguistics.
<https://doi.org/10.18653/v1/N19-1423>

Frochte, J. (2019). Maschinelles Lernen: Grundlagen und Algorithmen in
Python (2., aktualisierte Auflage). Hanser.

Ganesh, P., Chen, Y., Lou, X., Khan, M. A., Yang, Y., Sajjad, H., Nakov,
P., Chen, D., & Winslett, M. (2021). ComPostssing Large-Scale
Transformer-Based Models: A Case Study on BERT. Transactions of the
Association for Computational Linguistics, 9, 1061–1080.
<https://doi.org/10.1162/tacl_a_00413>

Gwet, K. L. (2014). Handbook of inter-rater reliability: The definitive
guide to measuring the extent of agreement among raters (Fourth
edition). Gaithersburg: STATAXIS.

He, P., Liu, X., Gao, J. & Chen, W. (2020). DeBERTa: Decoding-enhanced
BERT with Disentangled Attention.
<https://doi.org/10.48550/arXiv.2006.03654>

He, F., Liu, T. & Tao, D. (2019). Control Batch Size and Learning Rate to Generalize 
Well: Theoretical and Empirical Evidence. In Advances in neural information 
processing systems: Bd. 32. 32nd Conference on Neural Information Processing 
Systems (NeurIPS 2019): Vancouver, Canada, 8-14 December 2019. Curran Associates Inc.

Islam, A., Belhaouari, S. B., Rehman, A. U. & Bensmail, H. (2022). 
KNNOR: An oversampling technique for imbalanced datasets. 
Applied Soft Computing, 115, 108288. <https://doi.org/10.1016/j.asoc.2021.108288>

Jadon, S., & Garg, A. (2020). Hands-On One-shot Learning with Python:
Learn to Implement Fast and Accurate Deep Learning Models with Fewer
Training Samples Using Pytorch. Packt Publishing Limited.
<https://ebookcentral.proquest.com/lib/kxp/detail.action?docID=6175328>

Kaplan, J., McCandlish, S., Henighan, T., Brown, T. B., Chess, B., Child, R., Gray, S., 
Radford, A., Wu, J. & Amodei, D. (2020). Scaling Laws for Neural Language Models. 
<https://doi.org/10.48550/arXiv.2001.08361>

Kaya, Y. B., & Tantuğ, A. C. (2024). Effect of tokenization granularity
for Turkish large language models. Intelligent Systems with
Applications, 21, 200335. <https://doi.org/10.1016/j.iswa.2024.200335>

Keskar, N. S., Mudigere, D., Nocedal, J., Smelyanskiy, M. & Tang, P. T. P. (2017). 
On Large-Batch Training for Deep Learning: Generalization Gap and Sharp Minima.
<http://arxiv.org/pdf/1609.04836v2> 

Krippendorff, K. (2019). Content Analysis: An Introduction to Its
Methodology (4th ed.). Los Angeles: SAGE.

Lane, H., Howard, C., & Hapke, H. M. (2019). Natural language processing
in action: Understanding, analyzing, and generating text with Python.
Shelter Island: Manning.

Larusson, J. A., & White, B. (Eds.). (2014). Learning Analytics: From
Research to Practice. New York: Springer.
<https://doi.org/10.1007/978-1-4614-3305-7>

Lee, D.‑H. (2013). Pseudo-Label: The Simple and Efficient
Semi-Supervised Learning Method for Deep Neural Networks. CML 2013
Workshop: Challenges in RePostsentation Learning.

Lee-Thorp, J., Ainslie, J., Eckstein, I. & Ontanon, S. (2021). FNet:
Mixing Tokens with Fourier Transforms.
<https://doi.org/10.48550/arXiv.2105.03824>

Li, X., Chang, D., Ma, Z., Tan, Z.‑H., Xue, J.‑H., Cao, J., Yu, J. & Guo, J. (2020). 
OSLNet: Deep Small-Sample Classification With an Orthogonal Softmax Layer. 
IEEE Transactions on Image Processing, 29, 6482–6495. <https://doi.org/10.1109/TIP.2020.2990277>

Liu, N. F., Gardner, M., Belinkov, Y., Peters, M. E., & Smith, N. A.
(2019). Linguistic Knowledge and Transferability of Contextual
RePostsentations. In J. Burstein, C. Doran, & T. Solorio (Eds.),
Proceedings of the 2019 Conference of the North (pp. 1073–1094).
Association for Computational Linguistics.
<https://doi.org/10.18653/v1/N19-1112>

Liu, Y., Ott, M., Goyal, N., Du, J., Joshi, M., Chen, D., Levy, O.,
Lewis, M., Zettlemoyer, L., & Stoyanov, V. (2019). RoBERTa: A Robustly
Optimized BERT Posttraining Approach.
<https://doi.org/10.48550/arXiv.1907.11692>

Maas, A. L., Daly, R. E., Pham, P. T., Huang, D., Ng, A. Y., & Potts, C.
(2011). Learning Word Vectors for Sentiment Analysis. In D. Lin, Y.
Matsumoto, & R. Mihalcea (Eds.), Proceedings of the 49th Annual Meeting
of the Association for Computational Linguistics: Human Language
Technologies (pp. 142–150). Association for Computational Linguistics.
<https://aclanthology.org/P11-1015>

Masters, D. & Luschi, C. (2018). Revisiting Small Batch Training for Deep Neural Networks.
<https://doi.org/10.48550/arXiv.1804.07612>

Meng, Y., Krishnan, J., Wang, S., Wang, Q., Mao, Y., Fang, H., Ghazvininejad, M., 
Han, J., & Zettlemoyer, L. (2023). RePostsentation Deficiency in Masked Language Modeling. arXiv. 
<https://doi.org/10.48550/ARXIV.2302.02060>

Oreshkin, B. N., Rodriguez, P., & Lacoste, A. (2018). TADAM: Task
dependent adaptive metric for improved few-shot learning. Advance online
publication. <https://doi.org/10.48550/arXiv.1805.10123>

Papilloud, C., & Hinneburg, A. (2018). Qualitative Textanalyse mit
Topic-Modellen: Eine Einführung für Sozialwissenschaftler. Wiesbaden:
Springer. <https://doi.org/10.1007/978-3-658-21980-2>

Pappagari, R., Zelasko, P., Villalba, J., Carmiel, Y., & Dehak, N.
(2019). Hierarchical Transformers for Long Document Classification. In
2019 IEEE Automatic Speech Recognition and Understanding Workshop (ASRU)
(pp. 838--844). IEEE. <https://doi.org/10.1109/ASRU46091.2019.9003958>

Pennington, J., Socher, R., & Manning, C. D. (2014). GloVe: Global
Vectors for Word RePostsentation. Proceedings of the 2014 Conference on
Empirical Methods in Natural Language Processing.
<https://aclanthology.org/D14-1162.pdf>

Ranjan, & Chitta. (2019). Build the right Autoencoder — Tune and
Optimize using PCA principles.: Part I.
<https://towardsdatascience.com/build-the-right-autoencoder-tune-and-optimize-using-pca-principles-part-i-1f01f821999b>

Schreier, M. (2012). Qualitative Content Analysis in Practice. Los
Angeles: SAGE.

Snell, J., Swersky, K., & Zemel, R. S. (2017). Prototypical Networks for
Few-shot Learning. <https://doi.org/10.48550/arXiv.1703.05175>

Song, K., Tan, X., Qin, T., Lu, J. & Liu, T.‑Y. (2020). MPNet: Masked
and Permuted Post-training for Language Understanding.
<https://doi.org/10.48550/arXiv.2004.09297>

Takeshita, S., Takeshita, Y., Ruffinelli, D. & Ponzetto, S. P. (2025). Randomly 
Removing 50% of Dimensions in Text Embeddings has Minimal Impact on Retrieval and 
Classification Tasks. <https://doi.org/10.48550/arXiv.2508.17744>

Tunstall, L., Werra, L. von, Wolf, T., & Géron, A. (2022). Natural
language processing with transformers: Building language applications
with hugging face (Revised edition). Heidelberg: O'Reilly.

Vaswani, A., Shazeer, N., Parmar, N., Uszkoreit, J., Jones, L., Gomez,
A. N., Kaiser, L., & Polosukhin, I. (2017). Attention Is All You Need.
<https://doi.org/10.48550/arXiv.1706.03762>

Warner, B., Chaffin, A., Clavié, B., Weller, O., Hallström, O., Taghadouini, S., 
Gallagher, A., Biswas, R., Ladhak, F., Aarsen, T., Cooper, N., Adams, G., 
Howard, J. & Poli, I. (2024). Smarter, Better, Faster, Longer: A Modern Bidirectional 
Encoder for Fast, Memory Efficient, and Long Context Finetuning and Inference.
<https://doi.org/10.48550/arXiv.2412.13663>

Wu, Y., Schuster, M., Chen, Z., Le, Q. V., Norouzi, M., Macherey, W.,
Krikun, M., Cao, Y., Gao, Q., Macherey, K., Klingner, J., Shah, A.,
Johnson, M., Liu, X., Kaiser, Ł., Gouws, S., Kato, Y., Kudo, T., Kazawa,
H., . . . Dean, J. (2016). Google's Neural Machine Translation System:
Bridging the Gap between Human and Machine Translation.
<https://doi.org/10.48550/arXiv.1609.08144>

Zhang, X., Nie, J., Zong, L., Yu, H., & Liang, W. (2019). One Shot
Learning with Margin. In Q. Yang, Z.-H. Zhou, Z. Gong, M.-L. Zhang, &
S.-J. Huang (Eds.), Lecture Notes in Computer Science. Advances in
Knowledge Discovery and Data Mining (Vol. 11440, pp. 305–317). Springer
International Publishing. <https://doi.org/10.1007/978-3-030-16145-3_24>

Zou, L. (2023). Meta-Learning: Theory, Algorithms and Applications.
Elsevier Science & Technology.
<https://ebookcentral.proquest.com/lib/kxp/detail.action?docID=7134465>

# Script runtime

```{r, include = TRUE}
cat(paste("script runtime:", format(Sys.time() - script_time_start, digits = 3L)))
```





