---
title: "05 Sharing and Using Trained AI/Models"
author: "Florian Berding, Julia Pargmann, Andreas Slopinski, Elisabeth Riebenbauer, Karin Rebmann"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{05 Sharing and Using Trained AI/Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
running_num_figures=1
```

```{r setup, include = FALSE}
library(aifeducation)
```

# 1 Introduction

In the educational and social sciences, it is common practice to share
research instruments such as questionnaires or tests. For example, the
[Open Test Archive](https://www.testarchiv.eu/en) provides access to a
large number of open access instruments. *aifeducation* assumes AI-based
classifiers should be shareable, similarly to research instruments, to
empower educational and social science researchers and to support the
application of AI for educational purposes. Thus, *aifeducation* aims to
make the sharing process as convenient as possible.

For this aim, every model generated with *aifeducation* can be prepared
for publication in a few basic steps. In this vignette, we would like to
show you how to make your AI ready for publication. 

A explanation how to document a model with *AI for Education - Studio* can be 
found in the vignette [02 Using the graphical user interface Aifeducation -
Studio](https://fberding.github.io/aifeducation/articles/gui_aife_studio.html).
In this vignette we describe how to document a model with *R* syntax and provide some
hints for the content of documentation. 

# 2 Creating Documentation with *R* Syntax

The process for documenting a model with *R* syntax is similar for all models in
*aifeducation,* since all models use the same methods. Here, we will
illustrate the process for a `TextEmbeddingModel`.

First, every model needs a clear description of how it was developed,
modified and how it can be used. You can add a description via the
method `set_model_description`.

```{r, include = TRUE, eval=FALSE}
example_model$set_model_description(
  eng = NULL,
  native = NULL,
  abstract_eng = NULL,
  abstract_native = NULL,
  keywords_eng = NULL,
  keywords_native = NULL
)
```

This method allows you to provide a description in English and in the
native language of your model to make the distribution of your model
easier.

With `abstract_eng` and `abstract_native` you can provide a summary of
your description. This is very important if you would like to share your
work on a repository. With `keywords_eng` and `keywords_native` you can
set a vector of keywords, which helps others to find your work through
search engines.

You can access a model's description by using the method
`get_model_description`

```{r, include = TRUE, eval=FALSE}
example_model$get_model_description()
```

Besides a description of your work, it is necessary to provide
information on other people who were involved in creating the model.
This can be done with the method `set_publication_info`.

```{r, include = TRUE, eval=FALSE}
example_model$set_publication_info(
  type,
  authors,
  citation,
  url = NULL
)
```

First of all, you have to decide the type of information you would like
to add. You have two choices: "developer" and "modifier", which you set
with `type`.

-   `type="developer"` stores all information about the people involved
    in the process of developing the model. If you use a transformer
    model from [Hugging Face](https://huggingface.co/), the contributors
    and their description of the model should be entered as developers.
    In all other cases you can use this type for providing a description
    of how you developed the model.

-   In some cases you might wish to modify an existing model. This might
    be the case if you use a transformer model and you adapt the model
    to a specific domain or task. In this case you rely on the work of
    other people and modify their work. As such, you can describe your
    modifications by setting `type="modifier"`.

For every type of contributor you can add the relevant individuals via
`authors`. Please use the *R* function `personList()` for this. With
`citation` you can provide a short text on how to cite the work of the
different contributors. With `url` you can provide a link to relevant
sites of the model.

You can access the information by using `get_publication_info`.

```{r, include = TRUE, eval=FALSE}
example_model$get_publication_info()
```

Finally, you must provide a license for using your model. This can be
done with `set_model_license` and `get_model_license`.

```{r, include = TRUE, eval=FALSE}
example_model$set_model_license("GPL-3")
```

The documentation of your work is not part of the software. Here you can
set another license as for your software. You can set the license for your
documentation by using the method `set_documentation_license`.

```{r, include = TRUE, eval=FALSE}
example_model$set_documentation_license("CC BY-SA")
```

Now you are able to share your work. Please remember to save your now
fully described object as described in the vignette [03 Using R
syntax](classification_tasks.html).

The documentation process is the same for all models. There is only one difference. For
`TextEmbeddingModels`you can differentiate between "developers" and
"modifiers". This is not possible for the other models. For these models
you do not need the argument `type`. Calling this method would look
like:

```{r, include = TRUE, eval=FALSE}
example_model$set_publication_info(
  authors,
  citation,
  url = NULL
)
```

# 4 Content and Style of a Documentation

The necessary structure and content of a documentation depends on the
kind of model you would like to document, national laws (such as the
European AI Act), and the research standards of a discipline.

From a scientific point of view, we recommend that every model has an
abstract, keywords, and a detailed description in English. An additional
abstract, keywords, and a description in the native language of the
model may be helpful for reaching a broad audience in the corresponding
language community.

You can write your abstracts and descriptions in HTML and R Markdown
which allows you to add links to other sources or publications, to add
tables or to highlight important aspects of your model.

For all models we recommend that your description answers at least the
following questions:

-   Which kind of data was used to create the model?
-   How much data was used to create the model?
-   Which steps were performed and which method was used?
-   For which kinds of tasks or materials can the model be used?

This kind of information is necessary for others to form an opinion
about the model.

In the case of classifiers, we recommend to add some further
descriptions:

-   A short reference to the theoretical models that guided the
    development.
-   A clear and detailed description of every single category/class.
-   A short statement where the classifier can be used.
-   A description of the kind and quantity of data used for training.
-   Information on potential bias in the data.
-   If possible, information about the inter-coder-reliability of the
    coding process of the training data.
-   If possible, provide a link to the corresponding text embedding
    model or at least state where potential users can get the text
    embedding model.

The statement where to get the text embedding model is important since a
classifier can only be used with the corresponding text embedding model.

> Please do *not* report the performance values of your classifier in the
> description. These are displayed automatically in *AI for
> Education - Studio* or can be accessed directly via
> `example_classifier$reliability$test_metric_mean`.

